---
title: "Salary data department feature exploration"
author: "Mohammad Almerri"
date: "2024-11-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#clear existing environment objects
rm(list=ls())
library("tidyverse")
library("RnavGraphImageData")
library(purrr)
```

## The department feature

Loading the data:

```{r, echo=TRUE}
    
data <- tibble(read.csv("salary.csv"))

```

The department feature has the following amount of unique names:

```{r, echo=TRUE}

unique_departments <- unique(data$department)
print(length(unique_departments))
    
```

We need some way to lower the number of classes, either by merging or introducing a new column.

Starting with cleaning the dept names.

## Data proprocessing

### Convert to lower case

```{r, echo=TRUE}

data <- data %>%
  mutate(department_clean = tolower(department))
    
```

### Remove punctuaiton and special chars

```{r, echo=TRUE}

data <- data %>%
  mutate(department_clean = str_replace_all(department_clean, "[[:punct:]\\d]", ""))

    
```

### Remove punctuaiton and special chars

```{r, echo=TRUE}

data <- data %>%
  mutate(department_clean = str_replace_all(department_clean, "[[:punct:]\\d]", ""))

    
```

### Trim Whitespace

```{r, echo=TRUE}

data <- data %>%
  mutate(department_clean = str_squish(department_clean))

    
```
